---
title: "Akkermansia ASVs"
author: "Katherine Mueller"
date: "6/4/2021"
output: html_document
---

#Load Libraries
```{r}
library(phyloseq)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
```

#Define Path
```{r}
ms.rds <- "../../Data/Analysis/MS.rds"
out.dir <- "../../Data/Analysis"
```

#Load and Check Phyloseq Object
```{r}
ms.ps = read_rds(ms.rds)
print(ms.ps)
```

```{r}
sample_variables(ms.ps)
```

#Relative Abundance
```{r}
ms.rel.ps <- ms.ps %>%
  transform_sample_counts(function(x) x/sum(x))
```

#Akkermansia ASV Identification
```{r}
akks.ps <- ms.rel.ps %>%
  subset_taxa(Genus == "Akkermansia")
print(akks.ps)
```

Vizualize abundances of all four ASVs
```{r}
#Get dataframe for akk ASV abundances
all_akks <- akks.ps %>%
  otu_table() %>%
  as.data.frame()

#Make the dataframe long for plotting
akks.long <- all_akks %>%
  gather(ASV, Abundance)

#Plot Akk ASV abundances
akks.boxplot <- akks.long %>%
  ggplot(aes(x = ASV, y = Abundance))+
  geom_point() +
  geom_jitter()
akks.boxplot
```

Filter for Akk ASVs at a means abundance of at least 0.01%. Based on the above plot, this should probably result in three ASVs worth looking into.
```{r}
mean_cutoff = 0.0001
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

akks.0001mean.ps <- akks.ps %>%
  filter_taxa(mean_test, prune = TRUE)
print(akks.0001mean.ps)
```

I don't understand why the filter is returning an empty phyloseq object. It also returns empty even if the cutoff is set to zero.
```{r}
ASV_sequences <- akks.ps %>%
  otu_table() %>%
  as.data.frame() %>%
  colnames()
```

```{r}
ASV1.boxplot <- akks.long %>%
  filter(ASV == ASV_sequences[1]) %>%
  ggplot(aes(x = ASV, y = Abundance))+
  geom_point() +
  geom_jitter()
ASV2.boxplot <- akks.long %>%
  filter(ASV == ASV_sequences[2]) %>%
  ggplot(aes(x = ASV, y = Abundance))+
  geom_point() +
  geom_jitter()
ASV3.boxplot <- akks.long %>%
  filter(ASV == ASV_sequences[3]) %>%
  ggplot(aes(x = ASV, y = Abundance))+
  geom_point() +
  geom_jitter()
ASV4.boxplot <- akks.long %>%
  filter(ASV == ASV_sequences[4]) %>%
  ggplot(aes(x = ASV, y = Abundance))+
  geom_point() +
  geom_jitter()

plot_grid(ASV1.boxplot, ASV2.boxplot, ASV3.boxplot, ASV4.boxplot, labels = c(1,2,3,4))
```

It looks like the fourth ASV is the one that I expected to get rid of. I'll check mean abundances for all of them manually.
```{r}
mean(akks.long$Abundance[akks.long$ASV == ASV_sequences[1]])
head((akks.long$Abundance[akks.long$ASV == ASV_sequences[1]]))
```
```{r}
abundances1 <- (akks.long$Abundance[akks.long$ASV == ASV_sequences[1]])
typeof(abundances1)
abundances1
sum(abundances1)
```

Is it because of the one Nan?
```{r}
abundances1[is.na(abundances1)] <- 0
abundances1
mean(abundances1)
```

Taking out the NA fixed the mean issue, so I'll go ahead and double check the other three.
```{r}
abundances2 <- (akks.long$Abundance[akks.long$ASV == ASV_sequences[2]])
abundances2[is.na(abundances2)] <- 0
mean(abundances2)

abundances3 <- (akks.long$Abundance[akks.long$ASV == ASV_sequences[3]])
abundances3[is.na(abundances3)] <- 0
mean(abundances3)

abundances4 <- (akks.long$Abundance[akks.long$ASV == ASV_sequences[4]])
abundances4[is.na(abundances4)] <- 0
mean(abundances4)
```

My suspicion that the first three ASVs, but not the fourth, would pass the 0.01% filter was correct. I'm not sure how I could fix the NA issue in the phyloseq object to avoid this in the future though.
Get sequences
```{r}
akks0001 <- akks.ps %>%
  otu_table() %>%
  as.data.frame() %>%
  select(!ASV_sequences[4])
```

