---
title: "MS and Akkermansia Phylogroups"
author: "Katherine Mueller"
date: "6/8/2021"
output: html_document
---

#Load Libraries
```{r}
library(phyloseq)
library(cowplot)
library(car)
library(dplyr)
library(readr)
```

#Define Paths
```{r}
phylogroups.rds = "../../Data/Analysis/MS_wPhylogroups.rds"
out.dir = "../../Data/Analysis"
```

#Load and Check Phyloseq Object
```{r}
phylogroups.ps = read_rds(phylogroups.rds)
print(phylogroups.ps)
```

```{r}
sample_variables(phylogroups.ps)
```

##Analysis of Disease States
#Disease states
Determine which host disease states were recorded
```{r}
unique(sample_data(phylogroups.ps)$Host_disease)
```

Simplify disease state names
```{r}
sample_data(phylogroups.ps)$Host_disease[sample_data(phylogroups.ps)$Host_disease == "2RelapsingRemittingMS"] <- "RelapsingRemitting"
sample_data(phylogroups.ps)$Host_disease[sample_data(phylogroups.ps)$Host_disease == "1HealthyControl"] <- "Healthy"
sample_data(phylogroups.ps)$Host_disease[sample_data(phylogroups.ps)$Host_disease == "3SecondaryProgressiveMS"] <- "SecondaryProgressive"
sample_data(phylogroups.ps)$Host_disease[sample_data(phylogroups.ps)$Host_disease == "4PrimaryProgressiveMS"] <- "PrimaryProgressive"
```

```{r}
unique(sample_data(phylogroups.ps)$Host_disease)
```

```{r}
sample_data <- sample_data(phylogroups.ps) %>%
  as.matrix() %>%
  as.data.frame()
sample_data$Akk_abund <- as.numeric(sample_data$Akk_abund)
sample_data$AmI_abund <- as.numeric(sample_data$AmI_abund)
sample_data$AmII_abund <- as.numeric(sample_data$AmII_abund)
sample_data$AmIV_abund <- as.numeric(sample_data$AmIV_abund)
```

#Is ANOVA appropriate?
```{r}
akk_aov <- aov(Akk_abund ~ Host_disease,
               data = sample_data)

AmI_aov <- aov(AmI_abund ~ Host_disease,
               data = sample_data)

AmII_aov <- aov(AmII_abund ~ Host_disease,
                data = sample_data)

AmIV_aov <- aov(AmIV_abund ~ Host_disease,
                data = sample_data)

hist(akk_aov$residuals)
hist(AmI_aov$residuals)
hist(AmII_aov$residuals)
hist(AmIV_aov$residuals)
```

```{r}
qqPlot(akk_aov$residuals, id = FALSE)
qqPlot(AmI_aov$residuals, id = FALSE)
qqPlot(AmII_aov$residuals, id = FALSE)
qqPlot(AmIV_aov$residuals, id = FALSE)
```
Normality assumptions not met for ANOVA

#Non-parametric testing
```{r}
kruskal.test(Akk_abund ~ Host_disease, data = sample_data)

kruskal.test(AmI_abund ~ Host_disease, data = sample_data)

kruskal.test(AmII_abund ~ Host_disease, data = sample_data)

kruskal.test(AmIV_abund ~ Host_disease, data = sample_data)
```

No significant results for AmII and AmIV, but post-hoc pairwise comparisons can be used for all akks and AmI
```{r}
pairwise.wilcox.test(sample_data$Akk_abund, sample_data$Host_disease,
                     p.adjust.method = "BH")

pairwise.wilcox.test(sample_data$AmI_abund, sample_data$Host_disease,
                     p.adjust.method = "BH")
```

#Export dataframe for plotting in Prism
```{r}
write.csv(sample_data, file.path(out.dir, "MS_phylogroup_dataframe.csv"))
```

##Analysis of EDSS
